---
title: "Homework 4"
author: "Yuhan Xie"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    self_contained: yes
    toc: yes
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{The Homework 3 vignette}
-->
1. CASL Number 2 in Exercises 5.8.
The following is an example that $X^tX$ is well-conditioned, while the Hessian matrix is not due to a extreme vale of $\beta$, namely $p$ is close to 1. To make this vignette work, we set 'eval = F' for the second block. Otherwise, R will throw an error.
```{r}
n<-5000
p<-3
X<-cbind(1,matrix(rnorm(n*(p-1)),ncol = p-1))
XtX<-t(X)%*%X
solve(XtX)
```

```{r,eval=F}
beta<-c(10000000,200000000,100000000)
p<-1/(1+exp(-X%*%beta))
W<-as.numeric(p*(1-p))
H<-crossprod(X,diag(W)%*%X)
solve(H)
```

2. CASL Number 4 in Exercises 5.8.
Assume the distribution of y is from the exponential family
$$
f(y|\theta)=h(y)\exp\{\eta(\theta)T(y)-A(\theta)\}
$$

$$
\begin{aligned}
l^{pen}(y)&=l(y)-\lambda\|\boldsymbol{\beta}\|_2^2\\
&=\sum_{i=1}^n\{y_i\mathbf{x}_i^t\boldsymbol{\beta}-A(\mathbf{x}_i^t\boldsymbol{\beta})+\log h(y_i)\}-\lambda\boldsymbol{\beta}^T\boldsymbol{\beta}
\end{aligned}
$$
$$
\frac{\partial l^{pen}}{\partial\beta}=\frac{\partial l}{\partial \beta} - 2\lambda\boldsymbol{\beta}
$$

$$
\bigtriangledown_{\beta}l(y) = -X^t(y-\mathbb{E}y)-2\lambda\boldsymbol{\beta}
$$

$$
\frac{\partial^2l^{pen}}{\partial\beta\partial\beta^T}=\frac{\partial^2l}{\partial\beta\partial\beta^T}-2\lambda I_{p}
$$

$$
H(l)=-X^t\text{diag}(Var(y^{(k)})X-2\lambda I_{p}
$$

$$
\begin{aligned}
\beta^{(k+1)}&=\beta^{(k)}-H^{-1}(l)(\beta^{(k)})\bigtriangledown_{\beta}l(y)(\beta^{(k)})\\
&=\beta^{(k)}+[X^t\text{diag}(Var(y^{(k)}))X+2\lambda I_{p}]^{-1}(X^t(y-\mathbb{E}y^{(k)})-2\lambda\beta)\\
&=V^{-1}V\beta^{(k)}-2\lambda V^{-1}\beta^{(k)}+V^{-1}X^tWW^{-1}(y-\mathbb{E}y^{(k)})\\
&=V^{-1}[X^tWX+2\lambda I_{p}-2\lambda I_p]\beta^{(k)}+V^{-1}X^tWW^{-1}(y-\mathbb{E}y^{(k)})\\
&=V^{-1}X^tW\{X\beta^{(k)}+W^{-1}(y-\mathbb{E}y^{(k)})\}\\
&=[X^tWX+2\lambda I_{p}]^{-1}X^tWz
\end{aligned}
$$

$$
W=\text{diag}(Var(y^{(k)}))
$$

$$
V=X^tWX+2\lambda I_{p}
$$

$$
z=\{X\beta^{(k)}+W^{-1}(y-\mathbb{E}y^{(k)})\}
$$

```{r}
casl_glm_irwls_ridge <-
  function(X, y, family, maxit=25, tol=1e-10, lambda = 0){
    beta <- rep(0,ncol(X))
    for(j in seq_len(maxit))
    {
      b_old <- beta
      eta <- X %*% b_old
      mu <- family$linkinv(eta)
      mu_p <- family$mu.eta(eta)
      z <- eta + (y - mu) / mu_p
      W <- as.numeric(mu_p^2 / family$variance(mu))
      XtX <- crossprod(X, diag(W) %*% X) + 2*lambda*diag(1, ncol(X))
      Xtz <- crossprod(X, W * z)
      beta <- solve(XtX, Xtz)
      if(sqrt(crossprod(beta - b_old)) < tol) break
    }
    beta 
  }
```


3. Consider the sparse matrix implementation from class and the sparse add function:
```{r}
#Modification has been made to set an S4 class
sparse_add <- function(e1, e2){
            if(e1@dims[1] != e2@dims[1] || e1@dims[2] !=e2@dims[2]){
              stop("An error has occurred, please check your input!")
            }
              a <- data.frame(i = e1$i, j = e1$j, x = e1$x)
              b <- data.frame(i = e2$i, j = e2$j, x = e2$x)
              c <- merge(a, b, by = c("i", "j"), all = TRUE, suffixes = c("1", "2"))
              c$x1[is.na(c$x1)] <- 0
              c$x2[is.na(c$x2)] <- 0
              c$x <- c$x1 + c$x2
              c<-c[, c("i", "j", "x")]
              dims<-e1@dims
              return(sparse.matrix(c$i,c$j,c$x,dims = dims))
}
```

An example of function 'sparse_add'
```{r}
setClass("sparse.matrix", contains = "data.frame", slots = c(dims = "vector"))

sparse.matrix<-function(i,j,x,dims=NULL){
  if(is.null(dims)){
    dims<-c(max(i),max(j))
  }
  df<-data.frame(i = i, j = j , x = x)
  df<-df[order(df$i),]
  row.names(df)=NULL
  a1<-new("sparse.matrix",df,dims = dims)
  return(a1)
}
sm0 <- sparse.matrix(i = c(1, 2), j = c(1, 1), x = c(1, 1))
sparse_add(sm0,sm0)
```

-Implement a `sparse_multiply` function that multiplies two sparse matrices.
```{r}
#To implement the multiply function, we firstly implement a transpose function
transpose <- function(a) {
  dims<-c(a@dims[2],a@dims[1])
  temp<-a$i 
  a$i<-a$j
  a$j<-temp 
  a<-a[order(a$i),] 
  row.names(a)=NULL
  c<-sparse.matrix(a$i,a$j,a$x,dims)
  return(c)
}

#Multiplication function
sparse_multiply <-function(x, y){
  if(x@dims[2] != y@dims[1]){
    stop("An error has occurred, please check your input!")
  }
  dims<-c(x@dims[1],y@dims[2])
  a <- data.frame(i = x$i, j = x$j, x = x$x)
  # index arange of a
  ida<-numeric(a[nrow(a),1]+1)
  ida[1]<-1
  for(ra in 2:nrow(a)){
      if(a[ra-1,1]!=a[ra,1]){
        ida[a[ra,1]]<- ra
      }
  }
  ida[length(ida)] <- nrow(a)+1
  
  # index range of b
  y<-transpose(y)
  b<-data.frame(i = y$i, j = y$j, x = y$x)
  idb<-numeric(b[nrow(b),1]+1)
  idb[1]<-1
  for(rb in 2:nrow(b)){
    if(b[rb-1,1]!=b[rb,1]){
      idb[b[rb,1]]<- rb
    }
  }
  idb[length(idb)] <- nrow(b)+1
  
  df=data.frame()
  for(i in 1:(length(ida)-1)){
    tmp_a=a[ida[i]:(ida[i+1]-1),]
    for(j in 1:(length(idb)-1)){
      tmp_b=b[idb[j]:(idb[j+1]-1),]
      c <- merge(subset(tmp_a,select=c("j","x")), subset(tmp_b,select=c("j","x")), by = c("j"), all = TRUE, suffixes = c("1", "2"))
      c$x1[is.na(c$x1)] <- 0
      c$x2[is.na(c$x2)] <- 0
      row_sum <-  sum(c$x1 * c$x2)
      if (row_sum != 0)
      {
        df <- rbind(df, data.frame(i=ida[i],j=idb[j],x=row_sum))
      }
    }
  }
  row.names(df)=NULL
  sm<-sparse.matrix(df$i,df$j,df$x,dims)
  return(sm)
}
```

An example of function 'sparse_multiply'
```{r}
sm1 <- sparse.matrix(i = c(1, 2), j = c(1, 1), x = c(3, 1), dims = c(3, 2))
sm3 <- sparse.matrix(i = rep(1, 3), j = 1:3, x = 1:3, dims = c(2, 3))
sparse_multiply(sm1,sm3)
```

-Create a new class `sparse.matrix` that has add `+`, multiply `%*%`, and transpose `t()` methods.
```{r}
#Create a new class sparse.matrix
setClass("sparse.matrix", contains = "data.frame", slots = c(dims = "vector"))

sparse.matrix<-function(i,j,x,dims=NULL){
  if(is.null(dims)){
    dims<-c(max(i),max(j))
  }
  df<-data.frame(i = i, j = j , x = x)
  df<-df[order(df$i),]
  row.names(df)=NULL
  a1<-new("sparse.matrix",df,dims = dims)
  return(a1)
}

#Add transpose to the class
setMethod("t",signature="sparse.matrix",definition=function(x){
dims<-c(x@dims[2],x@dims[1])
temp<-x$i 
x$i<-x$j
x$j<-temp 
x<-x[order(x$i),] 
row.names(x)=NULL
c<-sparse.matrix(x$i,x$j,x$x,dims)
return(c)})

#Add "+" method to the class
setMethod("+",
          signature(e1= "sparse.matrix", e2 = "sparse.matrix"),
          function(e1, e2){
            if(e1@dims[1] != e2@dims[1] || e1@dims[2] ==e2@dims[2]){
              stop("An error has occurred, please check your input!")
            }
              a <- data.frame(i = e1$i, j = e1$j, x = e1$x)
              b <- data.frame(i = e2$i, j = e2$j, x = e2$x)
              c <- merge(a, b, by = c("i", "j"), all = TRUE, suffixes = c("1", "2"))
              c$x1[is.na(c$x1)] <- 0
              c$x2[is.na(c$x2)] <- 0
              c$x <- c$x1 + c$x2
              c<-c[, c("i", "j", "x")]
              dims<-e1@dims
              return(sparse.matrix(c$i,c$j,c$x,dims = dims))
          })

#Add %*% method to the class
setMethod("%*%",
          signature(x= "sparse.matrix", y = "sparse.matrix"),
          function(x, y){
            if(x@dims[2] != y@dims[1]){
              stop("An error has occurred, please check your input!")
            }
              dims<-c(x@dims[1],y@dims[2])
              a <- data.frame(i = x$i, j = x$j, x = x$x)
                # index arange of a
  ida<-numeric(a[nrow(a),1]+1)
  ida[1]<-1
  for(ra in 2:nrow(a)){
      if(a[ra-1,1]!=a[ra,1]){
        ida[a[ra,1]]<- ra
      }
  }
  ida[length(ida)] <- nrow(a)+1
  
  # index range of b
  y<-t(y)
  b<-data.frame(i = y$i, j = y$j, x = y$x)
  idb<-numeric(b[nrow(b),1]+1)
  idb[1]<-1
  for(rb in 2:nrow(b)){
    if(b[rb-1,1]!=b[rb,1]){
      idb[b[rb,1]]<- rb
    }
  }
  idb[length(idb)] <- nrow(b)+1
  
  df=data.frame()
  for(i in 1:(length(ida)-1)){
    tmp_a=a[ida[i]:(ida[i+1]-1),]
    for(j in 1:(length(idb)-1)){
      tmp_b=b[idb[j]:(idb[j+1]-1),]
      c <- merge(subset(tmp_a,select=c("j","x")), subset(tmp_b,select=c("j","x")), by = c("j"), all = TRUE, suffixes = c("1", "2"))
      c$x1[is.na(c$x1)] <- 0
      c$x2[is.na(c$x2)] <- 0
      row_sum <-  sum(c$x1 * c$x2)
      if (row_sum != 0)
      {
        df <- rbind(df, data.frame(i=ida[i],j=idb[j],x=row_sum))
      }
    }
  }
  row.names(df)=NULL
  sm<-sparse.matrix(df$i,df$j,df$x,dims)
  return(sm)
})
```

