---
title: "Homework 3"
author: "Yuhan Xie"
date: '`r Sys.Date()`'
output:
  pdf_document:
    toc: yes
  html_document:
    self_contained: yes
    toc: yes
  word_document:
    toc: yes
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{The Homework 3 vignette}
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(bis557)
```

#Page 117, Problem 7
```{r}
#kernel density function
kernel_epan <- function(x, h = 1) {
  x <- x/h
  ran <- as.numeric(abs(x) <= 1)
  val <- ((3/4)*(1-x^2)*ran)/h
  return(val)
}

kernel_density <- function(x, x_new, h) {
  sapply(x_new, function(v)
    { w <- kernel_epan(abs(x-v), h=h)
      f_den <- mean(w)
      f_den
    })
}
```

```{r}
set.seed(2018)
par(mfrow=c(3,1))
#Visualization
#bandwidth = 1.5,normal density N(0,1)
N <- 1000
x <- rnorm(N)
x_new <-seq(-5, 5, length.out = N)
bwidths<-1.5
f_den<-kernel_density(x,x_new,bwidths)
plot(x_new, f_den, xlab="x_new", ylab="kernel density", main="Normal Density Plot",type="l",ylim = c(0,0.6),lwd = 2)
true_den <- dnorm(x_new, 0, 1)
lines(x_new, true_den, col="green",lwd = 2)
legend("topright", legend = c("Kernel Density","True Density"), col=c("black", "green"),lty=c(1,1))

#bandwidth = 0.5,normal density N(0,1)
N <- 1000
x <- rnorm(N)
x_new <-seq(-5, 5, length.out = N)
bwidths<-0.5
f_den<-kernel_density(x,x_new,bwidths)
plot(x_new, f_den, xlab="x_new", ylab="kernel density", main="Normal Density",type="l",ylim = c(0,0.6),lwd = 2)
true_den <- dnorm(x_new, 0, 1)
lines(x_new, true_den, col="green",lwd = 2)
legend("topright", legend = c("Kernel Density","True Density"), col=c("black", "green"),lty=c(1,1))

#bandwidth = 0.1,normal density N(0,1)
N <- 1000
x <- rnorm(N)
x_new <-seq(-5, 5, length.out = N)
bwidths<-0.1
f_den<-kernel_density(x,x_new,bwidths)
plot(x_new, f_den, xlab="x_new", ylab="kernel density", main="Normal Density",type="l",ylim = c(0,0.6))
true_den <- dnorm(x_new, 0, 1)
lines(x_new, true_den, col="green",lwd = 2)
legend("topright", legend = c("Kernel Density","True Density"), col=c("black", "green"),lty=c(1,1))
```

```{r}
set.seed(2018)
par(mfrow=c(3,1))
#bandwidth = 1.5, chi-square density(df=3)
N <- 1000
x <- rchisq(N,df=3)
x_new <-seq(0, 10, length.out = N)
true_den <- dchisq(x_new,df=3)
bwidths<-1.5
f_den<-kernel_density(x,x_new,bwidths)
plot(x_new, f_den, xlab="x_new", ylab="kernel density", main="Chi-square Density",type="l",ylim = c(0,0.3),lwd = 2)
lines(x_new, true_den, col="purple",lwd = 2)
legend("topright", legend = c("Kernel Density","True Density"), col=c("black", "purple"),lty=c(1,1))

#bandwidth = 0.5, chi-square density(df=3)
N <- 1000
x <- rchisq(N,df=3)
x_new <-seq(0, 10, length.out = N)
true_den <- dchisq(x_new,df=3)
bwidths<-0.5
f_den<-kernel_density(x,x_new,bwidths)
plot(x_new, f_den, xlab="x_new", ylab="kernel density", main="Chi-square Density",type="l",ylim = c(0,0.3),lwd = 2)
lines(x_new, true_den, col="purple",lwd = 2)
legend("topright", legend = c("Kernel Density","True Density"), col=c("black", "purple"),lty=c(1,1))

#bandwidth = 0.1, chi-square density(df=3)
N <- 1000
x <- rchisq(N,df=3)
x_new <-seq(0, 10, length.out = N)
true_den <- dchisq(x_new,df=3)
bwidths<-0.1
f_den<-kernel_density(x,x_new,bwidths)
plot(x_new, f_den, xlab="x_new", ylab="kernel density", main="Chi-square Density Plot",type="l",ylim = c(0,0.3),lwd = 2)
lines(x_new, true_den, col="purple",lwd = 2)
legend("topright", legend = c("Kernel Density","True Density"), col=c("black", "purple"),lty=c(1,1))
```

```{r,eval=FALSE}
#a sequence of bandwidth
set.seed(2018)
N <- 1000
x <- rnorm(N)
x_new <-seq(-10, 10, length.out = N)
y_new <- dnorm(x_new)
mses<-rep(0,N)
f_den<-rep(0,N)
bwidths<-seq(0.1, 5, length.out = N)
for (k in 1:N) {
  f_den<-kernel_density(x = x,x_new = x_new,h=bwidths[k])
  mses[k]<-mean((f_den-y_new)^2)
}
best_h<-which.min(mses)
sprintf("mse: %01.09f h= %01.04f",mses[best_h],bwidths[best_h])
plot(bwidths,mses,main = "MSEs Vs Bandwidths")
```


#Page 200,Problem 3
$f:\mathbb{R}^p\rightarrow \mathbb{R},g:\mathbb{R}^p\rightarrow \mathbb{R}$
For any values $b_1,b_2\in \mathbb{R}^p$ and quantity $t\in[0,1]$
Denote:
$$
(f+g)(x)=f(x)+g(x)
$$
\begin{align*}
(f+g)(tb_1+(1-t)b_2)&=f(tb_1+(1-t)b_2)+g(tb_1+(1-t)b_2)\\
&\leq tf(b_1)+(1-t)f(b_2)+tg(b_1)+(1-t)g(b_2)\\
&=t(f+g)(b_1)+(1-t)(f+g)(b_2)
\end{align*}

Thus, $f+g$ is a convex funciton.

#Page 200,Problem 4
For any values $b_1,b_2\in \mathbb{R}^p$ and quantity $t\in[0,1]$
\begin{align*}
|tb_1+(1-t)b_2|&\leq|tb_1|+|(1-t)b_2|\\
&\leq t|b_1|+(1-t)|b_2|
\end{align*}
Thus, the absolute value function is convex.

$$
||v||_1=\sum_j|v_j|
$$
where $v_j$ is the jth element in $v$.
As $|v_j|$ is convex, from Problem 3 we know that the sum of convex function is also convex. Thus, the $l_1-norm$ is also convex.

#Page 200,Problem 5
$l_2-norm$
$$
||v||_2=\sum_jv_j^2
$$
where $v_j$ is the jth element in $v$.

For any values $b_1,b_2\in \mathbb{R}^p$ and quantity $t\in[0,1]$
$$
t^2b_1^2+(1-t)^2b_2^2+2t(1-t)b_1b_2-tb_1^2-(1-t)b_2^2=t(t-1)(b_1-b_2)^2\leq0
$$
$$
(tb_1+(1-t)b_2)^2=t^2b_1^2+(1-t)^2b_2^2+2t(1-t)b_1b_2\leq tb_1^2+(1-t)b_2^2
$$

$f:\mathbb{R}^p\rightarrow \mathbb{R}$, f is a convex function.
For any $a>0$, 
$$
af(tb_1+(1-t)b_2)\leq taf(b_1)+(1-t)af(b_2)
$$
Thus, $af$ is convex.

Thus, the square function is a convex function. Then, the sum of square functions ($l_2-norm$) is also convex. As elastic net objective function is the sum of $l_1$ norm and $l_2$ norms, it is convex.

#Page 200,Problem 6
We create the check KKT function with formula from (7.37) on page 188
$$
\frac{1}{n}\sum_{i=1}^nx_{il}(y_i-\sum_{j=1}^px_{ij}\widehat{b_j})=\lambda s_l
$$
$$
s_j\in\left\{\begin{matrix}
 1& if\quad \widehat{b}_j>0\\ 
 -1& if\quad \widehat{b}_j<0\\ 
 [-1,1]& if\quad \widehat{b}_j=0
\end{matrix}\right.
$$

```{r}
#The function to test if the KKT condition is violated
lasso_reg_with_screening<-function(X,y,b,lambda){
  resids<-y-X%*%b
  s<-apply(X, 2, function(xj) crossprod(xj,resids))/nrow(X) 
  s<-round(s/lambda)
  #round s to make it the same as in cv.glmnet function
  ret_vec<-rep(TRUE,length(b))
  for (i in 1:length(b)) {
    if(b[i]>0)
     ret_vec[i]<- !(s[i] == 1) 
    if(b[i]==0)
     ret_vec[i]<- !(abs(s[i])<=1)
    if(b[i]<0) 
        ret_vec[i]<- !(s[i] == -1)}
  #if KKT condition is violated, we return a vector contains at least one TRUE
  #if KKT condition is not violated, we return a vector of FALSEs
  return(ret_vec)
}
```

```{r}
library(glmnet)
X<-model.matrix(mpg~.-1,data=mtcars[,1:7])
X<-scale(X)
y<-mtcars$mpg
y<-y-mean(y)
glnnet_fit<-cv.glmnet(X,y,standardize = FALSE,intercept = FALSE)
lambda<-glnnet_fit$lambda.1se
lambda
b<-glnnet_fit$glmnet.fit$beta[,(glnnet_fit$lambda)==lambda]
b
#Every beta complies the KKT condition given lambda, as we get a vector of FALSEs
lasso_reg_with_screening(X,y,b,lambda)
```

```{r}
#Manually change the coefficient of 'cyl' to 0
#Violation of KKT exists as there are Trues
b[1]<-0
lasso_reg_with_screening(X,y,b,lambda)
```

```{r}
#Manually change the coefficient of 'disp' to be not equal to 0
#Violation of KKT exists as there are Trues
b<-glnnet_fit$glmnet.fit$beta
b<-b[,(glnnet_fit$lambda)==lambda]
b[2]<-1
lasso_reg_with_screening(X,y,b,lambda)
```