---
title: "Homework 2"
author: "Yuhan Xie"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: yes
    toc: true
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{The Homework 2 vignette}
-->

## The ridge_reg function

Ridge regression is a technique for analyzing multivariate regression data that suffer from multicollinearity. 

```{r}
library(bis557)
data("ridge_train")
data("ridge_test")
```

The function to calculate mean square error of the ridge regression model.
```{r}
mse_ridge_residuals <- function(formula,lambda){ 
  #Get X and y from the test dataset
  y_test<- model.frame(formula,ridge_test)[,1]
  X_test<- model.matrix(formula,ridge_test)
  #Fit the ridge regression model with the train dataset
  fit_ridge<-ridge_reg(formula,lambda,ridge_train)
  betas<-fit_ridge$coef
  #Calculate the estimated y value
  y_hat<-X_test%*%betas
  #Calculate the mean square error
  mean((y_hat-y_test)^2)
}
```

Plot lambda on a log scale with its corresponding mean square error.
```{r}
lambda<-10^(seq(-2,5,0.05))
mse<-sapply(lambda, function(x) mse_ridge_residuals(y ~. ,x))
plot(log(lambda),mse)
```

Find the optimal value of lambda and the corresponding minimum value of mean square error.
```{r}
f<-function(x) mse_ridge_residuals(y ~. ,x)
lambda_optimal<-optimize(f,interval = c(-100,1000000))
cat("The optimal value of lambda is",lambda_optimal$minimum,"\n")
cat("The minimum value of mse is",lambda_optimal$objective)
```

