---
title: "Homework 3"
author: "Yuhan Xie"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: yes
    toc: true
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{The Homework 3 vignette}
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(bis557)
```


#Page 200,Problem 3
$f:\mathbb{R}^p\rightarrow \mathbb{R},g:\mathbb{R}^p\rightarrow \mathbb{R}$
For any values $b_1,b_2\in \mathbb{R}^p$ and quantity $t\in[0,1]$
Denote:
$$
(f+g)(x)=f(x)+g(x)
$$
\begin{align*}
(f+g)(tb_1+(1-t)b_2)&=f(tb_1+(1-t)b_2)+g(tb_1+(1-t)b_2)\\
&\leq tf(b_1)+(1-t)f(b_2)+tg(b_1)+(1-t)g(b_2)\\
&=t(f+g)(b_1)+(1-t)(f+g)(b_2)
\end{align*}

Thus, $f+g$ is a convex funciton.

#Page 200,Problem 4
For any values $b_1,b_2\in \mathbb{R}^p$ and quantity $t\in[0,1]$
\begin{align*}
|tb_1+(1-t)b_2|&\leq|tb_1|+|(1-t)b_2|\\
&\leq t|b_1|+(1-t)|b_2|
\end{align*}
Thus, the absolute value function is convex.

$$
||v||_1=\sum_j|v_j|
$$
where $v_j$ is the jth element in $v$.
As $|v_j|$ is convex, from Problem 3 we know that the sum of convex function is also convex. Thus, the $l_1-norm$ is also convex.

#Page 200,Problem 5
$l_2-norm$
$$
||v||_2=\sum_jv_j^2
$$
where $v_j$ is the jth element in $v$.

$$
t^2b_1^2+(1-t)^2b_2^2+2t(1-t)b_1b_2-tb_1^2-(1-t)b_2^2=t(t-1)(b_1-b_2)^2\leq0
$$
$$
(tb_1+(1-t)b_2)^2=t^2b_1^2+(1-t)^2b_2^2+2t(1-t)b_1b_2\leq tb_1^2+(1-t)b_2^2
$$
Thus, the square function is a convex function. Then, the sum of squares function ($l_2-norm$) is also convex. As elastic net objective function is the sum of $l_1$ norm and $l_2$ norms, it is convex.

#Page 200,Problem 6
We create the check KKT function with formula from (7.37) on page 188
$$
\frac{1}{n}\sum_{i=1}^nx_{il}(y_i-\sum_{j=1}^nx_{ij}\widehat{b_j})=\lambda s_l
$$
$$
s_j\in
$$

```{r}
#The function to test if the KKT condition is violated
lasso_reg_with_screening<-function(X,y,b,lambda){
  resids<-y-X%*%b
  s<-apply(X, 2, function(xj) crossprod(xj,resids))/nrow(X) 
  s<-round(s/lambda)
  #round s to make it the same as in cv.glmnet function
  ret_vec<-rep(FALSE,length(b))
  for (i in 1:length(b)) {
    if(b[i]>0)
     ret_vec[i]<- !(s[i] == 1) 
    if(b[i]==0)
     ret_vec[i]<- !(abs(s[i])<=1)
    if(b[i]<0) 
        ret_vec[i]<- !(s[i] == -1)}
  #if KKT condition is violated, we return a vector contains at least one TRUE
  #if KKT condition is not violated, we return a vector of FALSEs
  return(ret_vec)
}
```

```{r}
library(glmnet)
X<-model.matrix(Sepal.Length~.-1,data=iris)
X<-scale(X)
y<-iris$Sepal.Length
y<-y-mean(y)
glnnet_fit<-cv.glmnet(X,y,standardize = FALSE,intercept = FALSE)
lambda<-glnnet_fit$lambda.1se
lambda
b<-glnnet_fit$glmnet.fit$beta[,(glnnet_fit$lambda)==lambda]
#After checking betas, we find the coefficient of 'Speciesversicolor' is equal to 0, all other coefficients are not equal to 0.
b
#Every beta complies the KKT condition given lambda, as we get a vector of FALSEs
lasso_reg_with_screening(X,y,b,lambda)
```

```{r}
#Manually change the coefficient of 'Sepal.Width' to 0
#Violation of KKT exists as there are Trues
b[1]<-0
lasso_reg_with_screening(X,y,b,lambda)
```

```{r}
#Manually change the coefficient of 'Speciesversicolor' to not be 0
#Violation of KKT exists as there are Trues
b<-glnnet_fit$glmnet.fit$beta[,(glnnet_fit$lambda)==lambda]
b[5]<-1
lasso_reg_with_screening(X,y,b,lambda)
```